All Your Graphs Belong to Us

### 2012-07-19 doc/asplos13/grappa.tex
Rework of "Grappa" section

In this section we describe the Grappa runtime system (Figure~\ref{fig:grappa}).  Grappa is a software runtime system that one adds to a commodity HPC cluster.  It provides the following software abstractions to applications: \textbf{memory}, a global distributed shared memory and \textbf{tasking}, co-operative multithreading with automatic load balancing.  Under the hood a third component, \textbf{communication}, a network packet scheduler/aggregator is key to achieving high performance, but largely invisible to applications.  We will now describe each of these components.

\subsection{Memory}

@@

Local memory is exactly what its name implies.  Accesses occur through conventional pointers.  The compiler emits an access and the memory is manipulated directly.  Obviously, local pointers cannot access memory on other system nodes.  Applications use local accesses for a number of things in Grappa: the stack associated with a task, accesses to localized global memory in caches (see below), and accesses to debugging infrastructure that is local to each system node.

Global memory is a distributed block cyclic heap.  Applications allocate memory in this heap and a global pointer is returned.  Global pointers are utilized in one of two ways: (1) via a \emph{cache} or (2) via a \emph{delegate}.

\paragraph{Cache:} With a cache, applications can instruct Grappa to fetch a global pointer (of any length) and return a local pointer to a cached copy of the global memory.  Under the hood, Grappa performs the mechanics of gathering chunks of data from multiple system nodes and presenting a conventional appearing linear block of memory as a pointer into a cache.  Caches are either read-only or read-write and alternatively incoherent or coherent.  Read-only and and read-write caches are what their names imply.  Under the hood, the distinction between these two forms of caches lays in how they interact with the directories that maintain the coherence state for coherent access to the memory the global pointer points to.  For coherent caches, these directories maintain a simple coherence protocol that permits multiple outstanding read-only cached copies of data to exist, or alternatively, only a single outstanding read-write cached copy to exist.  Unlike hardware caches, these software caches require explicit release operations issued in software; i.e., applications must explicitly inform Grappa when the local pointer that points into a local cached copy of global memory is no longer needed.  Grappa then uses this release operation to interact with the directories that control the global memory in question.  For incoherent accesses, the directories are bypassed.  Applications acquire a local pointer to a cached copy of the global data and can expressly push the data from this local copy back to global memory in order to post any writes.  Obviously, applications should only use incoherent accesses when they can \emph{prove} accesses across multiple threads and system nodes are disjoint, or solely read-only, or (in rarer cases) maintaining a consistent view of memory is irrelevant to application correctness, or for debugging purposes.

\paragraph{Delegates:} Alternatively applications can dispatch computation to be performed on individual machine-word sized chunks of global memory to the memory system itself (e.g., \emph{fetch-and-add}).  Applications provide a pointer to a function and a global pointer to Grappa, which takes care to execute that function on the system node that actually stores the memory pointed to by the global pointer.  In such cases, localization of the pointer is more efficient.  Delegate operations are most often used for synchronization primitives, but really any computation is permissible.

\TODO{Check: Do delegate operations interact with the directories?  Are they coherent?}

Figure~\ref{fig:globalmemory} depicts the global memory architecture and access interface.

\subsection{Tasking}

\subsection{Communication}

With Grappa, from the perspective of the application, all communication between tasks occurs through global shared memory.  These requests are ultimately translated into explicit inter system node communication requests that are dispatched to GASnet.  Under the hood there exists a key component within Grappa, the message aggregator, that makes fine-grained communication efficient on commodity Infiniband networks.  Infiniband achieves its peek bisection bandwidth \emph{only} when the packet sizes are relatively large (a kilobyte or more).  In our experience, graph algorithms request objects through the caching system (previously described) with a median object size of \checkme{16} bytes.  Our measurements, confirm manufactures published data~\cite{infinibandbandwidth}, that with this packet size the bisection bandwidth is only a small fraction, less than ~\checkme{5\%} of the peak bisection bandwidth.  The reason for this discrepancy is the combination of overheads associated with handling each packet (in terms of bytes that form the actual Infiniband packet, processing time at the card and processing on the CPU within the driver stack).

To work around this problem we perform packet aggregation.  Instead of dispatching a packet immediately to the Infiniband hardware, Grappa holds onto an individual communication request.  Pending requests that are to be sent to the same remote node are queued.  Once a sufficient number of them (in terms of bytes) or a timeout occurs, they are dispatched as a single packet to GASnet and onward to the Infiniband network.  In Section~\ref{sec:results} we explore the sensitivity of performance to these two key parameters, but the net result is that performance is maximized by aiming to aggregate for \checkme{1024} bytes and \checkme{1ms}.

>>>>

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.5\columnwidth]{figs/grappa-crappy}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} Grappa runtime system \vspace{-4ex}}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

In this section we describe the Grappa runtime system
(Figure~\ref{fig:grappa}). Grappa is a software runtime system that
allows a commodity distributed-memory HPC cluster to be programmed as
if it were a single large shared-memory machine. Grappa is designed to
smooth over some of the performance discontinuities in commodity
hardware, giving good performance when there is little locality to be
exploited while allowing the programmer to exploit it when it is
available.

Grappa provides three key capabilities. The first is a distributed
shared \textbf{global memory} with high aggregate random access
bandwidth. The second is a \textbf{tasking} system with lightweight
multithreading and automatic load balancing through work
stealing. Finally, under the hood, a \textbf{communication} layer uses
network packet scheduling and aggregation to achieve high performance
while remaining largely invisible to applications.

In the following sections, we discuss the programmer's view of
Grappa's main capabilities, including a subset of the Grappa API. In
section~\ref{implementation}, we discuss the details of their
implementation.

\subsection{Tasks}

%The basic unit of locality in Grappa is the {\em core}. Each core is
%responsible for a section of the global memory

The basic unit of execution in Grappa is the {\em task}. Tasks are
small; just a function pointer and some arguments. Tasks are only
queued when they are spawned; later, when resources are free, they are
allocated a stack, bound to a core, and executed.

What makes Grappa's tasks special is that when they perform
long-latency operations they can give up control of their core,
allowing the core to remain busy while waiting for the operation to
complete. This most often done implicitly inside a call into Grappa's
API, but it can also be done explicitly by the programmer using the
calls shown in Figure~\ref{fig:scheduling}.

All Grappa scheduling is done in userspace, and minimal state is
saved, in order to minimize the context switch time. In our
experiments, we saw context switch times as low as \TODO{40ns}.

This lightweight context switching is Grappa's key enabling
feature. It gives us the ability to tolerate latency. Given that
ability, we are able to trade latency for throughput: by {\em
  increasing} latency in key components in the system we are able to
increase our aggregate random access bandwidth, our synchronization
bandwidth, and our ability to tolerate load.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
    \item[ \texttt{ yield() } ] \hfill \\
      Gives up control of core to scheduler, queuing task to be scheduled again soon
    \item[ \texttt{ suspend() } ] \hfill \\
      Gives up control of core to scheduler
    \item[ \texttt{ wake( task * $t$ ) } ] \hfill \\
      Enqueues $t$ to be scheduled again soon
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:scheduling} Grappa API: scheduling} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

\subsection{Expressing parallelism}

A programmer's goal in coding with Grappa should be to express as much
parallelism as possible without worrying about where it will execute.
Grappa then chooses where and when to execute this parallelism,
scheduling as much work as necessary to tolerate network latencies and
keep the systems' cores busy.

Grappa provides multiple methods for expressing parallelism, shown in
Figure~\ref{fig:expressing-parallelism}. The first is explicit task
spawns. When the programmer identifies work that can be done in
parallel, the work may be wrapped up in a function and queued with its
arguments for later execution using a \texttt{spawn}. Sometimes the
programmer may want to spawn a task on a specific core in the system,
or at the home core of a particular memory location. Grappa provides a
\texttt{spawn\_on} call for this purpose; the task is bound the
specified core immediately and will never be stolen.

The other method for expressing parallelism is a Cilk-like {\em
  recursive loop decomposition} \TODO{cite}. The programmer specifies
a function pointer along with start and end indices and an optional
threshold; Grappa decomposes the iterations into a
logarithmically-sized tree of tasks, stopping to execute when the
number of iterations is below the threshold.

Finally, a programmer may want to run a small piece of code on a
particular core in the system without waiting for execution resources
to be available. Grappa provides the \texttt{call\_on} call for this
purpose.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
    \item[ \texttt{spawn( void (*fp)(args) )} ] \hfill \\
      Creates a new stealable task
    \item[ \texttt{spawn\_on( core, (*fp)(args) )} ] \hfill \\
      Creates a new private task that will run on a specific core 
    \item[ \texttt{parallel\_for( (*fp)(args), start, end )} ] \hfill \\
      Executes iterations of a loop as stealable tasks 
    \item[ \texttt{call\_on( core, (*fp)(args) )} ] \hfill \\ 
      Runs a limited function on a specific core without consuming
      execution resources
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:expressing-parallelism} Grappa API: expressing parallelism} % \vspace{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

\subsection{Accessing memory}

@@

Local memory is local to a single core in the system.  Accesses occur
through conventional pointers.  The compiler emits an access and the
memory is manipulated directly.  Applications use local accesses for a
number of things in Grappa: the stack associated with a task, accesses
to localized global memory in caches (see below), and accesses to
debugging infrastructure that is local to each system node.  Local
pointers cannot access memory on other cores, and are valid only on
their home core.

Large, low-locality data is stored in Grappa's global memory. All
global data must be accessed through calls into Grappa's API, shown in
Figure~\ref{fig:accessing-memory}.

Grappa provides two methods for storing data in the global memory. The
first is a distributed heap striped across all the machines in the
system. The \texttt{global\_malloc} and \texttt{global\_free} calls
are used to allocate and deallocate memory in the global heap; on
allocation, a global pointer is returned. Grappa also allows any local
data on the core's stacks or heap to be exported to the global address
space and accessed by other cores across the system.

There are two approaches to accessing global memory. When the
programmer expects global data to have spatial locality to exploit,
{\em cache} operations may be used. When there is no locality to
exploit, {\em delegate} operations are used. Since these operations
are expected communicate with other nodes and have high latency,

With a cache, applications can instruct Grappa to fetch a global
pointer of any length and return a local pointer to a cached copy of
the global memory. Under the hood, Grappa performs the mechanics of
gathering chunks of data from multiple system nodes and presenting a
conventional appearing linear block of memory as a pointer into a
cache. Grappa cache operations have the usual read-only and read-write
variants, along with a write-only variant used to initialize data
structures.

When accessing and modifying low-locality data, it is more efficient
to modify the data on its home core rather than bringing a copy to the
requesting core and returning it after modification. Delegate
operations provide this capability. Applications can dispatch
computation to be performed on individual machine-word sized chunks of
global memory to the memory system itself (e.g.,
\emph{fetch-and-add}).

Delegate operations are also the primary synchronization method in
Grappa. Each piece of global memory is paired with a single core in
the system; all accesses to that memory are done by that core. This
allows delegate operations to be performed atomically on the core
simply by making sure no other code interleaves; since context
switches in Grappa are cooperative, this is trivial. Thus, delegate
operations are able to provide atomic semantics without using atomic
operations.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
      \item[ \texttt{ global\_address global\_malloc( size )} ] \hfill \\
      \item[ \texttt{ global\_free( global\_address )} ] \hfill \\
        Allocates and frees memory in the global heap
      \item[ \texttt{ delegate\_read( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_write( global\_address, local\_var )} ] \vspace{-2ex}
      \item[ \texttt{ delegate\_cas( global\_address, local\_var )} ] \vspace{-2ex}
      \item[ \texttt{ delegate\_inc( global\_address, local\_var )} ] \vspace{-2ex} \hfill \\
        Performs a memory operation at the home core of a global address
      \item[ \texttt{ cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} ]
      \item[ \texttt{ cache\_release( global\_address, local\_buf )} ] \vspace{-2ex} \hfill \\
        Perform cache operations \TODO{expand}
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:accessing-memory} Grappa API: accessing memory} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

\subsection{Under the hood}

Grappa contains two key components that are not directly visible to
the programmer but are key to providing good performance. Both depend
on Grappa's ability to tolerate latency in order to obtain this
performance.

The first is a load balancing mechanism that uses work stealing to
move tasks from busy nodes to idle nodes. When a core detects that it
has idle execution resources and no new tasks to run, it issues steal
requests over the network. The core continues to execute remaining
active work while the steal requests are in flight. When a steal
request succeeds, it returns with more tasks that can be matched with
execution resources and scheduled.

The second is a network packet scheduling and aggregation system. From
the perspective of the application, all communication between tasks in
Grappa takes place either through global shared memory or remote task
spawns. Both of these requests are ultimately translated into explicit
inter-system-node/core communication requests that are dispatched to
the Garnet library. Most of our requests are small, and commodity
networks do not transport small packets efficiently. In order to make
more efficient use of bandwidth, Grappa aggregates packets going to the
same destination until it can form a larger packet that will make more
efficient use of the network. In this way, Grappa trades per-request
latency for application-level throughput.

### 2012-07-19 doc/asplos13/implementation.tex
Rework of "Grappa" section

\paragraph{2D global addresses:} All global addresses in Grappa are 64-bit values. Each byte of global
memory is associated with a {\em home core}; this core is responsible
for allocating and modifying that memory. Since today's commodity
processors support a 48-bit virtual address space \TODO{cite?}, Grappa
uses the spare bits to distinguish which address space a global
address belongs to, and to represent the home core of a global
address.

@@

Any local data can be made accessible by other nodes in the system by
wrapping the address and node ID into a 2D global address. This
address can then be accessed with delegate or cache operations. At the
destination the address is converted into a canonical x86 address by
replacing the upper bits with the sign-exteded upper bit of the
virtual address.

\paragraph{Linear global addresses:} The second global address space is the {\em linear} address space, so
named because nearby addresses can be treated as adjacent even if they
are stored on different nodes. Addresses in the linear address space
are partitioned across the cluster in a block-cyclic fashion. Choosing
the block size involves trading off sequential bandwidth against
aggregate random access bandwidth. Smaller block sizes help spread
data across all the memory controllers in the cluster, but larger
block sizes allow the locality-optimized memory controllers to provide
increased sequential bandwidth. The block size, which is configurable, is typically set to 64
bytes, or the size of a single hardware cache line, in order to
exploit spatial locality when available.

@@

All global memory operations are implemented with active
messages. Requestors construct a descriptor holding the current
task/worker ID and space for a response, issue their request active
message to the node given by the request address, and suspend
themselves. When the request arrives at the address' home node, the
operation is performed and a response active message is sent. When the
response arrives at the requesting node, the descriptor is filled in
with the results of the operation and the requesting task is woken.

\paragraph{Delegates:} Delegate operations are used for syncronization. Their modifications
are always executed at the home core of their address. Each task may
have only one outstanding delegate operation. We limit delegate
operations to operate on objects in the 2D address space or objects that
fit in a single block of the linear address space. No context switches
are allowed while the data it being modified. Given these
restrictions, we can ensure that delegate operations for the same address from multiple
requestors are always serialized through a single core in the system,
providing atomic semantics without using atomic operations.

Cache operations are used to exploit spatial or temporal
locality. They allow a single task to buffer a range of data, read or
modify it, and write the modified version back to its home node. All
acquires and releases are expressed explicitly by the
programmer. Cache operations may operate on data that spans multiple
blocks of the linear address space; the programmer provides the
starting address and length, and the cache issues as many active
messages as necessary to retrieve all necessary data.

@@

All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfereing
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.}

@@

Grappa tasks are 32-byte entities: a 64-bit function pointer plus
three 64-bit arguments. We use three arguments because tasks are
usually generated as part of a parallel loop decomposition, and thus
each task needs three kinds of data:
\begin{description}
\item[Function pointer] This 64-bit value indicates what routine
  should run. We configure the system nodes to disable address randomization, so that we can depend on the fact that function pointers are valid across process images.
\item[Private argument] This 64-bit value is typically used for task-specific
  data (e.g., a loop index).
\item[Shared argument] This 64-bit value is typically used for data shared
  between all tasks that are part of a loop.
\item[Synchronization argument] This 64-bit value is typically used to determine
  when all tasks that are part of a loop have finished. It is usually
  a global pointer to a synchronization object allocated at the core
  that spawned the task.
\end{description}
While these are the most common uses of the three task arguments, they
are treated as arbitray 64-bit values in the runtime, and can be used
for any purpose.

@@

Context switching in Grappa is non-preemptive. Any Grappa routine that
does a potentially-long-latency operation calls a context-switch
routine that gives control of the core back to the scheduler. Users
may also explicity yield control to the scheduler.

@@

\TODO{What is a worker?  We don't realyl describe this.  -M}

@@

Most study of work stealing assumes that a processor has one worker
thread and so it only steals when utilization would go to zero. 
Grappa, on the other hand, relies on having many worker threads per core. The
core would be fully-utilized if there is always something on the ready
queue. Even if there are many active tasks, if they are all suspended
for long-latency network requests, then the core is underutilized.
So, there is a choice: should the core use some of this idle time to
perform work stealing or should it just wait as it is possible that local tasks will
create more work. We choose to have a core inititiate a steal when all
of the following conditions hold: no workers are ready to run, 
the unstarted task queues are empty, and there are no outstanding steal requests. 
Having only one outstanding steal request throttles steals to
prevent flooding the network, but other scalable quieting mechanisms
would be possible, such as a voting tree\cite{scalableWorkStealingOrCilk98}
or lifelines \cite{lifelines}. Termination detection is not
built into the Grappa task scheduler, rather, it is considered a programming error for
the program to exit without syncing all tasks.

\TODO{where to put the following related}
Fork join computations: implement with futures or continuation passing
style
feed-forward computations: implement with scalable task counting (we
focus on this style in our benchmarks)
(breifly describe global task joiner)

@@

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. In our applications we see that the bulk of
the messages sent are small. \TODO{include stats} Commodity networks,
even ones designed for high-performance computing, provide low
bandwidth utilization at small message sizes. \TODO{include stats?} To
make the best use of the network, we must convert our small messages
into large ones.

@@

There are three situations in which a queue of aggregated messages is
sent. First, each queue has a message size threshold, chosen to give
reasonable network performance. If the size in bytes of a queue is
above the threshold, the contents of the queue are sent
immediately. Second, each queue has a wait time threshhold. If the
oldest message in a queue has been waiting longer than this threshold,
the contents of the queue are sent immediately, even if the queue size
is lower than the message size threshold.  Third, queues may be
explicitly flushed in situations where the programmer wants to
minimize the latency of a message at the cost of bandwidth
utilization.

>>>>

\paragraph{2D global addresses} All global addresses in Grappa are 64-bit values. Each byte of global
memory is associated with a {\em home core}; this core is responsible
for allocating and modifying that memory. Since today's commodity
processors support a 48-bit virtual address space \TODO{cite?}, Grappa
uses the spare bits to distinguish which address space a global
address belongs to, and to represent the home core of a global
address.

@@

Any local data can be made accessible by other nodes in the system by
wrapping the address and node ID into a 2D global address. This
address can then be accessed with delegate or cache operations. At the
destination the address is converted into a canonical x86 address by
replacing the upper bits with the sign-extended upper bit of the
virtual address.

\paragraph{Linear global addresses} The second global address space is the {\em linear} address space, so
named because nearby addresses can be treated as adjacent even if they
are stored on different nodes. Addresses in the linear address space
are partitioned across the cluster in a block-cyclic fashion. Choosing
the block size involves trading off sequential bandwidth against
aggregate random access bandwidth. Smaller block sizes help spread
data across all the memory controllers in the cluster, but larger
block sizes allow the locality-optimized memory controllers to provide
increased sequential bandwidth. The block size, which is configurable, is typically set to 64
bytes, or the size of a single hardware cache line, in order to
exploit spatial locality when available.

@@

All global memory operations are implemented with active
messages. Requesters construct a descriptor holding the current
task/worker ID and space for a response, issue their request active
message to the node given by the request address, and suspend
themselves. When the request arrives at the address' home node, the
operation is performed and a response active message is sent. When the
response arrives at the requesting node, the descriptor is filled in
with the results of the operation and the requesting task is woken.

\paragraph{Delegates} Delegate operations are used for low-locality accesses and synchronization. Their modifications
are always executed at the home core of their address. Each task may
have only one outstanding delegate operation. We limit delegate
operations to operate on objects in the 2D address space or objects that
fit in a single block of the linear address space. No context switches
are allowed while the data it being modified. Given these
restrictions, we can ensure that delegate operations for the same address from multiple
requesters are always serialized through a single core in the system,
providing atomic semantics without using atomic operations.

\paragraph{Caches} Cache operations are used to exploit spatial or temporal
locality. They allow a single task to buffer a range of data, read or
modify it, and write the modified version back to its home node. All
acquires and releases are expressed explicitly by the
programmer. Cache operations may operate on data that spans multiple
blocks of the linear address space; the programmer provides the
starting address and length, and the cache issues as many active
messages as necessary to retrieve all necessary data.

@@

All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfering
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.}

@@

Grappa tasks are 32-byte entities: a 64-bit function pointer plus
three 64-bit arguments. We use three arguments because tasks are
usually generated as part of a parallel loop decomposition, and thus
each task needs three kinds of data:
\begin{description}
\item[Function pointer] This 64-bit value indicates what routine
  should run. We configure the system nodes to disable address randomization, so that we can depend on the fact that function pointers are valid across process images.
\item[Private argument] This 64-bit value is typically used for task-specific
  data (e.g., a loop index).
\item[Shared argument] This 64-bit value is typically used for data shared
  between all tasks that are part of a loop.
\item[Synchronization argument] This 64-bit value is typically used to determine
  when all tasks that are part of a loop have finished. It is usually
  a global pointer to a synchronization object allocated at the core
  that spawned the task.
\end{description}
While these are the most common uses of the three task arguments, they
are treated as arbitrary 64-bit values in the runtime, and can be used
for any purpose.

@@

Context switching in Grappa is non-preemptive. Any Grappa routine that
does a potentially-long-latency operation calls a context-switch
routine that gives control of the core back to the scheduler. Users
may also explicit yield control to the scheduler.

@@

\TODO{What is a worker?  We don't really describe this.  -M}

@@

Most study of work stealing assumes that a processor has one worker
thread and so it only steals when utilization would go to zero. 
Grappa, on the other hand, relies on having many worker threads per core. The
core would be fully-utilized if there is always something on the ready
queue. Even if there are many active tasks, if they are all suspended
for long-latency network requests, then the core is underutilized.
So, there is a choice: should the core use some of this idle time to
perform work stealing or should it just wait as it is possible that local tasks will
create more work. We choose to have a core initiate a steal when all
of the following conditions hold: no workers are ready to run, 
the unstarted task queues are empty, and there are no outstanding steal requests. 
Having only one outstanding steal request throttles steals to
prevent flooding the network, but other scalable quieting mechanisms
would be possible, such as a voting tree\cite{scalableWorkStealingOrCilk98}
or lifelines \cite{lifelines}. Termination detection is not
built into the Grappa task scheduler, rather, it is considered a programming error for
the program to exit without syncing all tasks.

\TODO{where to put the following related}
Fork join computations: implement with futures or continuation passing
style
feed-forward computations: implement with scalable task counting (we
focus on this style in our benchmarks)
(briefly describe global task joiner)

@@

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. Commodity networks including Infiniband
achieves their peak injection bandwidth \emph{only} when the packet
sizes are relatively large---on the order of multiple kilobytes. The
reason for this discrepancy is the combination of overheads associated
with handling each packet (in terms of bytes that form the actual
Infiniband packet, processing time at the card and processing on the
CPU within the driver stack). Our measurements confirm manufactures
published data~\cite{infinibandbandwidth}, that with this packet size
the bisection bandwidth is only a small fraction, less than
~\checkme{5\%} of the peak bisection bandwidth.

In our experiments the vast majority of requests were smaller than 255
bytes \TODO{more precise? median?}, far too small to make efficient
use of the network. To make the best use of the network, we must
convert our small messages into large ones.

@@

There are three situations in which a queue of aggregated messages is
sent. First, each queue has a message size threshold, chosen to give
reasonable network performance. If the size in bytes of a queue is
above the threshold, the contents of the queue are sent
immediately. Second, each queue has a wait time threshold. If the
oldest message in a queue has been waiting longer than this threshold,
the contents of the queue are sent immediately, even if the queue size
is lower than the message size threshold.  Third, queues may be
explicitly flushed in situations where the programmer wants to
minimize the latency of a message at the cost of bandwidth
utilization.
